---
title: "Homework 1"
author: "Mayuri Albal"
date: "2022-09-18"
output: github_document
---
```{r setup, include = FALSE}
library(tidyverse)
```
# Problem 1
### Data Loading
```{r}
data("penguins", package = "palmerpenguins")
```

### Data Summary
```{r}
summary(penguins)
```

### Observations
```{r}
nrow(penguins)
ncol(penguins)
```
In this data set, there are 8 variables  and 344 observations for penguins. 
The variables include:species, island, length of bill (mm), depth of bill (mm),
flipper length (mm), body mass (g), sex, and year of birth. 

### Mean Flipper Length
```{r}
mean.default(penguins$flipper_length_mm, na.rm = TRUE)
```
The mean flipper length of penguins in the data set is 200.92 mm.

### Plots
```{r}
ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, color= species))+    geom_point(na.rm = TRUE)
ggsave("penguin_plot.pdf", height = 4, width = 6)
```

# Problem 2

### Create Data Frame
```{r dataframe_creation}
icecream_df= tibble(
  icecream= rnorm(10),
  icecream_logical= icecream > 0,
  icecream_char = c("vanilla", "chocolate", "strawberry", "mintcc", "cookiescream", "rockyroad", 
                     "birthdaycake", "coffee", "butterpecan", "saltedcaramel"),
  icecream_factor = factor(c("base", "base", "addin", "newflavor", "addin", "addin",
                              "addin", "newflavor", "addin", "newflavor"))
  )
icecream_df
```
### Mean of Each Variable in DF
```{r}
mean(pull(icecream_df,icecream))
mean(pull(icecream_df,icecream_logical))
mean(pull(icecream_df,icecream_char))
mean(pull(icecream_df,icecream_factor))
```
### Reasoning
In this new data frame, variables that are numeric are able to have a mean value generated
for them, as observed with ice cream due to it having a value between 0 and 1. A mean was generated for the 'icecream' and 'icecream_logical' variables. However, due to the other variables being character variables with no value assigned to them, a mean
value cannot be generated.

### Numeric Function
```{r}
as.numeric(pull(icecream_df, icecream_logical))
as.numeric(pull(icecream_df, icecream_char))
as.numeric(pull(icecream_df, icecream_factor))
```
When using the numeric function for each of the variables, it is evident why a mean
value could not be taken. For the logical vector, the values outputted were 0 and 1 
signifying the true and false response generated, however a mean value cannot be taken 
of these responses as it is not possible to have a middle value for true or false in
the frame. For the character vector, all responses were NA due to the character nature 
thus with no numeric value they cannot have a generated mean. Lastly, for the factor
vector, the output was the levels from 1 to 3, however similar to the logic vectors, there
cannot be a mean response of the three levels created, as each of these level values
are also arbitrarily assigned values. As a result a mean cannot be taken for these variables
due to the returned values being uninterpretable in the context of the data frame.